On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-02-21 MS.tex
Change prediction for RGR vs AGR in results section

By definition, plants with high SLA can deploy a unit of leaf area more cheaply. However, high-SLA species also have shorter leaf lifespan, and must use more energy replacing leaves lost via turnover \citep{Wright:2004jb}. These two effects occur within the terms  $\frac{\ud B}{\ud t}$ (net live biomass production) and $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ (marginal cost of deploying additional leaf area), appearing on the right side in the equations for growth rate (\ref{eq:eq1}, \ref{eq:eq2}, \ref{eq:eq3}). Whether plant growth rate increases with SLA depends on how the relative magnitude of the benefit (cheap leaf construction, via $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ ) and cost (higher leaf turnover, via $\frac{\ud B}{\ud t}$) of increased SLA changes with size (Table \ref{tab:trade-offs}).
During early life, when leaves comprise a large part of the plant, increasing SLA has an overwhelmingly positive effect on height and diameter growth rate, because the effect of reducing $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ (the cost of producing a unit of leaf area) has a large effect compared to the other terms in eq. \ref{eq:daldmt}. As plants increase in size, however, the cost of building sapwood ($\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$) increases disproportionately, decreasing the benefit of cheaper leaf construction in eq.  \ref{eq:daldmt}. Consequently, the effect of leaf turnover becomes relatively stronger at larger sizes, and as such, the effect of SLA on height, diameter and, mass growth shifts from positive to negative as plants increase in size. Note in the case of mass growth, the effect of cheaper leaf construction is realised not in the $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ term , but as a greater total leaf area at a given mass, i.e. the accumulated effect of past reductions in  $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$.

@@

The only effect of increased seed size in the above model is to produce a larger overall seedling. Under similar light conditions, larger seedlings are therefore predicted to have faster absolute growth rates. However, because the ratio of leaf area to total biomass decreases with plant size, larger seedlings are predicted to have lower mass-based growth rates during the earliest phases of growth \citep{Turnbull:2012ew}. Yet, as plants grow, we expect the effect of this initial size difference to be minimal, and the correlation between seed size and growth rate should therefore disappear. For larger plants, seed mass is not expected to have any distinct influence on growth rates  (Table \ref{tab:trade-offs}).

@@

We had hypotheses for seed mass, SLA and wood density, about a change in the correlation between trait and growth rate if absolute growth rate was used instead of relative growth rate, or if mass was used instead of diameter and height. For seed mass, effect size differed significantly between relative growth rate and absolute growth rate at seedling stage; as expected the correlation with relative growth rate was positive and shifted to non significant with absolute growth rate (Fig \ref{fig:fig4}d; CI not overlapping). For SLA, the correlation was expected to shift from positive to non significant if growth in mass was measured on absolute instead of relative terms; however, none of the data were measured with absolute rates in the seedling stage (Fig. \ref{fig:figA8}). For wood density, some differences occurred but they were not easily attributed to any single cause. In particular, the correlation at juvenile stage shifted from negative with relative growth rate to non significant with absolute growth rate (i.e CI included zero, Fig. \ref{fig:fig4}b). Yet, the correlation coefficients recorded for absolute growth rate came from \citet{Augspurger:1984ct}, and unlike others they used wood density values from adult plants together with growth from juveniles. So it was not possible to disentangle the effect of growth measurement from the effect of trait origin. For other traits, effect sizes were generally consistent between relative growth rate and absolute growth rate when both measurements were available.

>>>>

By definition, plants with high SLA can deploy a unit of leaf area more cheaply. However, high-SLA species also have shorter leaf lifespan, and must use more energy replacing leaves lost via turnover \citep{Wright:2004jb}. These two effects occur within the terms  $\frac{\ud B}{\ud t}$ (net live biomass production) and $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ (marginal cost of deploying additional leaf area), appearing on the right side in the equations for growth rate (\ref{eq:eq1}, \ref{eq:eq2}, \ref{eq:eq3}). Whether plant growth rate increases with SLA depends on how the relative magnitude of the benefit (cheap leaf construction, via $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ ) and cost (higher leaf turnover, via $\frac{\ud B}{\ud t}$) of increased SLA changes with size (Table \ref{tab:trade-offs}).
During early life, when leaves comprise a large part of the plant, increasing SLA has an overwhelmingly positive effect on height and diameter growth rate, because the effect of reducing $\frac{\ud M_{\rm l}}{\ud A_{\rm l}}$ (the cost of producing a unit of leaf area) has a large effect compared to the other terms in eq. \ref{eq:daldmt}. As plants increase in size, however, the cost of building sapwood ($\frac{\ud M_{\rm s}}{\ud A_{\rm l}}$) increases disproportionately, decreasing the benefit of cheaper leaf construction in eq.  \ref{eq:daldmt}. Consequently, the effect of leaf turnover becomes relatively stronger at larger sizes, and as such, the effect of SLA on height, diameter and, mass growth shifts from positive to negative as plants increase in size (Table \ref{tab:trade-offs}). Note in the case of mass growth, the effect of cheaper leaf construction is realised not in the $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$ term , but as a greater total leaf area at a given mass, i.e. the accumulated effect of past reductions in  $\frac{\ud A_{\rm l}}{\ud M_{\rm a} }$.

@@

The only effect of increased seed size in the above model is to produce a larger overall seedling. Under similar light conditions, larger seedlings are therefore predicted to have faster absolute growth rates . However, because the ratio of leaf area to total biomass decreases with plant size, larger seedlings are predicted to have lower mass-based relative growth rates during the earliest phases of growth \citep{Turnbull:2012ew}(Table \ref{tab:trade-offs}). Yet, as plants grow, we expect the effect of this initial size difference to be minimal, and the correlation between seed size and growth rate should therefore disappear. For larger plants, seed mass is not expected to have any distinct influence on growth rates (Table \ref{tab:trade-offs}).

@@

We had hypotheses for seed mass about a change in the correlation between trait and growth rate if absolute growth rate was used instead of relative growth rate. Effect size for seed mass differed significantly between relative growth rate and absolute growth rate at seedling stage; as expected the correlation with relative growth rate was positive and shifted to non significant with absolute growth rate (Fig \ref{fig:fig4}d; CI not overlapping). For wood density, some differences occurred according to the growth measurement used but they were not easily attributed to any single cause. In particular, the correlation at juvenile stage shifted from negative with relative growth rate to non significant with absolute growth rate (i.e CI included zero, Fig. \ref{fig:fig4}b). Yet, the correlation coefficients recorded for absolute growth rate came from \citet{Augspurger:1984ct}, and unlike others they used wood density values from adult plants together with growth from juveniles. So it was not possible to disentangle the effect of growth measurement from the effect of trait origin. For other traits, effect sizes were generally consistent between relative growth rate and absolute growth rate when both measurements were available.
